---
title: "Causal Impact in R with Advertising Data"
output:
  html_document:
    keep_md: true
    toc: true
    toc_depth: 3
    toc_float: true
    toc_collapsed: true
  pdf_document:
      toc: true
      number_sections: true
  theme: lumen
---
# Introduction

This R Markdown document serves as a guide to introduce users to demonstrating impact of Ads in R with a real example used with Advertising data. 

Consider: https://github.com/klarsen1/MarketMatching

# Required packages
```{r}
library(CausalImpact)
library(tidyverse)
library(xts)
```

# Import data files

```{r}
ad_impressions <- read_csv('DV360 ad impressions 20180901-20181231.csv')

head(ad_impressions)

```

```{r}
channel_grouping <- read_csv('Channel Grouping Sessions.csv')
head(channel_grouping)
```

Problem Statement: We want to know the impact of K-Addict campaign impressions on the number of sessions on K Bright website

Variables of interest in an experiment (those that are measured or observed) are called response or dependent variables. Other variables in the experiment that affect the response and can be set or measured by the experimenter are called predictor, explanatory, or independent variables. 

For example, you might want to determine the recommended baking time for a cake recipe. 

Predictor variables such as Baking time or oven temperature would affect the response (we would change or optimise these), which could be moisture level or thickness of the cake (the outcome that we would want to measure).

Datasets are obtained from Google Analytics, from 1 September 2018 to 31 December 2018
Response: date, Sessions (1 August 2018 - 31 December 2018) 
First group of users (Predictor 1): date, Impressions of Flaunting Divas campaign (1 Sept 2018 - 31 Oct 2018)
Second group of users (Predictor 2): date, Impressions of K-Bright Visitor campaign (optional)

http://www.dartistics.com/timeseries.html

Channels report: ChannelGrouping, Date, Sessions
Ad Impressions report: DV360 Exchange (CM Model), DV360 Creative Name (CM Model), Date, DV360 Impressions

```{r}
# Create Date-Time column
ad_impressions$Date <-  as.Date(as.character(ad_impressions$Date),format="%Y%m%d")
#ad_impressions$Date <-  as.POSIXct(ad_impressions$Date, format = "%Y%m/%d", #origin="SGT")

channel_grouping$Date <-  as.Date(as.character(channel_grouping$Date),format="%Y%m%d")
#channel_grouping$Date <-  as.POSIXct(channel_grouping$Date, format = "%Y%m/%d", #origin="SGT")

# Filter for the specific campaign that we want to test
ad_impressions_K_Addict <- ad_impressions %>% 
  filter(`DV360 Exchange (CM Model)` == 'DoubleClick Ad Exchange' &
           `DV360 Creative Name (CM Model)` == 'Flaunting Divas') %>% 
  select(-`DV360 Exchange (CM Model)`)

# remove unwanted columns and join datasets
channel_grouping_wide <- channel_grouping %>% 
  spread(`Default Channel Grouping`, Sessions)

ad_impressions_K_Addict_wide <- ad_impressions_K_Addict %>% 
  spread(`DV360 Creative Name (CM Model)`, `DV360 Impressions`)

channel_ad_impressions <- channel_grouping_wide %>% 
  left_join(ad_impressions_K_Addict_wide, by = 'Date') %>% 
   mutate_at(vars(-Date), funs(replace(., is.na(.), 0)))

channel_ad_impressions_xts <- xts(channel_ad_impressions[-1], order.by = as.Date(channel_ad_impressions$Date))

## data in order of response, predictor1, predictor2, etc.
model_data <- channel_ad_impressions_xts[,c("Display","Flaunting Divas","Paid Search")]

```

# Create Pre & Post Periods

```{r}
# Determine pre and post event time frames varying post period timepre.period 
pre.period <- as.Date(c("2018-08-01","2018-08-31"))
post.period <- as.Date(c("2018-11-01","2018-12-31"))
```


# Run Causal Impact analysis and see results
```{r}
impact_channel_ad_impressions_xts = CausalImpact(model_data, pre.period, post.period)
summary(impact_channel_ad_impressions_xts)

```
# Generate a narrated report of the results
```{r}
summary(impact_channel_ad_impressions_xts, "report")
```

# Plot graphs of the results
```{r}
plot(impact_channel_ad_impressions_xts)
```
Original: Solid, black line is observed data before the intervention and dotted, blue line is the predicted coin in values for what would have occurred without the intervention

Pointwise: The net difference between the observed and predicted response on the original scale, or the difference between the solid, black line and the dotted, blue line on the original graph.

Cumulative: Dotted, blue line is the individual causal effects added up in time, day after day.

For all three graphs, the light blue shaded area represents the results in a 95% confidence level. The farther that the graph extends past the beginning of the intervention, the less certain of the causal effect; hence, the larger the shaded area.

# Show the predictor variables used in the model

```{r}
plot(impact_channel_ad_impressions_xts$model$bsts.model, "coefficients")
```

# Further exploration

https://ee.cooper.edu/~keene/assets/Neema_Thesis_vFinal.pdf