---
title: "Analytics Checkup"
output:
  pdf_document: default
  html_document:
    df_print: paged
    latex_engine: xelatex
---
# to add: pages per session, % of returning users, sales by channelgrouping

```{r setup, include=FALSE, echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo=FALSE, cache=TRUE, error=TRUE, message=FALSE, warning = FALSE, fig.width=10)
options(scipen = 666)
options(httr_oauth_cache = FALSE)
# google analytics libs
library(googleAnalyticsR)
#library(googleAuthR)
library(skimr)      # for summary table
library(knitr)      # to kable
library(kableExtra) # nice kables
library(tidyverse)  # data wrangling
library(viridis)    # colors
library(glue)       # paste++
library(lubridate)  # date funcs
library(gridExtra)  # line up graphs

customer <- 'SQ'
accountid <- '17015991'       # ga account id
property <- 'UA-17015991-1'   # ga property id
viewid <- '46948678'          # ga view id

# authenticate - interactive. Tip: use a service account + json file
ga_auth(token = ".httr-oauth")
# look back 28 days
from_date <- Sys.Date() - 28 
to_date   <- Sys.Date() - 1
date_range <- c(as.character(from_date), as.character(to_date))

basic_metrics <- c("users", "sessions", "pageViews", 
                   "totalEvents", "transactions", "transactionRevenue",
                   "bounceRate", "pageviewsPerSession",
                   "transactionsPerSession", "newUsers")

basic_metrics_edit <- c("users", "sessions", "pageViews", 
                   "totalEvents", "transactions", "transactionRevenue",
                   "bounceRate", "pageviewsPerSession",
                   "transactionsPerSession", "returningUsers")
   
basics <-  google_analytics(viewid, 
                            date_range = date_range,
                            dimensions = c("date"), 
                            metrics = basic_metrics)

basics <- basics %>% 
  mutate(returningUsers = users - newUsers) %>% 
  select(-newUsers)

channel_metrics <- c("sessions")
   
channel_basics <-  google_analytics(viewid, 
                                    date_range = date_range,
                                    dimensions = c("date", 
                                                   "channelGrouping"),
                                    metrics = channel_metrics)
channel_basics_spread <- channel_basics %>% 
  mutate(channelGrouping = gsub(" ", "_", channelGrouping)) %>%
  mutate(channelGrouping = gsub("\\(", "", channelGrouping)) %>% 
  mutate(channelGrouping = gsub(")", "", channelGrouping)) %>% 
  spread(channelGrouping, sessions)

conversion_metrics <- c("transactions", "sessions")

conversion_basics <-  google_analytics(viewid, 
                                       date_range = date_range,
                                       dimensions = c("date",
                                                      "channelGrouping",
                                                      "deviceCategory"),
                                       metrics = conversion_metrics)

conversion_basics_spread <- conversion_basics %>% 
  mutate(channelGrouping = gsub(" ", "_", channelGrouping)) %>%
  mutate(channelGrouping = gsub("\\(", "", channelGrouping)) %>% 
  mutate(channelGrouping = gsub(")", "", channelGrouping)) %>%
  group_by(date, channelGrouping) %>% 
  summarise(conversionRate = round(100*(sum(transactions)/sum(sessions)), 2)) %>% 
  spread(channelGrouping, conversionRate)

conversion_device_basics_spread <- conversion_basics %>% 
  mutate(channelGrouping = gsub(" ", "_", channelGrouping)) %>%
  mutate(channelGrouping = gsub("\\(", "", channelGrouping)) %>% 
  mutate(channelGrouping = gsub(")", "", channelGrouping)) %>%
  group_by(date, deviceCategory) %>% 
  summarise(conversionRate = round(100*(sum(transactions)/sum(sessions)), 2)) %>% 
  spread(deviceCategory, conversionRate)


productCat_metrics <- c("itemRevenue")

productCat_basics <-  google_analytics(viewid, 
                                       date_range = date_range,
                                       dimensions = c("date",
                                                      "productName"),
                                       metrics = productCat_metrics)
# find top 10 products by product revenue
pdt_summary <- productCat_basics %>% 
  group_by(productName) %>% 
  summarise(itemRevenue = sum(itemRevenue)) %>% 
  top_n(10) %>% 
  select(-itemRevenue) %>% 
  pull(productName)


productCat_basics <- productCat_basics %>% 
  filter(productName %in% pdt_summary)

productCat_basics_spread <- productCat_basics %>%
  mutate(productName = gsub(" ", "_", productName)) %>% 
  mutate(productName = gsub("\\(", "", productName)) %>% 
  mutate(productName = gsub(")", "", productName)) %>% 
  mutate(productName = gsub("\\|", "_", productName)) %>% 
  mutate(productName = gsub("\\/", "_", productName)) %>% 
  spread(productName, itemRevenue)

cdims <- ga_custom_vars_list(accountid, property, type = "customDimensions")
cmets <- ga_custom_vars_list(accountid, property, type = "customMetrics")
```

This is an automatically generated document, generated on `r Sys.Date()` - for view `r viewid` - **`r customer`**

Used date range is `r from_date` to `r to_date`

## Eyeball Check

Checking some basic metrics to see how it looks.

```{r}
plot_metric_by_day <- function(metric) {
  ggplot(basics, aes(x = date, color = "value")) +
    geom_line(aes_string(y = metric)) +
    theme_minimal() + xlab("") + ylab("value") +
    ggtitle(glue("28 day trend of {metric}")) +
    theme(plot.title = element_text(size = 10)) +
    scale_color_viridis(discrete = TRUE) + theme(legend.position = "none")
    
}
plotlist <- list()
for (m in basic_metrics_edit) {
  plotlist[[m]] <- plot_metric_by_day(m)
}
# arrange plots in a grid
do.call("grid.arrange", c(plotlist, ncol=floor(sqrt(length(plotlist)))))
```

Showing a summary and percentiles:

```{r echo=FALSE}
skim_to_wide(basics[2:ncol(basics)]) %>%
  select(variable, mean, sd, p0, p50, p100, hist)
```


Checking Channel Grouping metrics by session to see how it looks.

```{r}
channel_list <- unique(channel_basics$channelGrouping)
channel_list <- gsub(" ", "_", channel_list)
channel_list <- gsub("\\(", "", channel_list)
channel_list <- gsub(")", "", channel_list)

plot_metric_by_day <- function(metric) {
  ggplot(channel_basics_spread, aes(x = date, color = "value")) +
    geom_line(aes_string(y = metric)) +
    theme_minimal() + xlab("") + ylab("value") +
    ggtitle(glue("28 day trend of {metric}")) +
    theme(plot.title = element_text(size = 10)) +
    scale_color_viridis(discrete = TRUE) + theme(legend.position = "none")
    
}
plotlist <- list()
for (m in channel_list) {
  plotlist[[m]] <- plot_metric_by_day(m)
}
# arrange plots in a grid
do.call("grid.arrange", c(plotlist, ncol=floor(sqrt(length(plotlist)))))
```

Showing a summary and percentiles:

```{r}
skim_to_wide(channel_basics_spread[2:ncol(channel_basics_spread)]) %>%
  select(variable, mean, sd, p0, p50, p100, hist)
```



Checking trends for sales metrics for each  product category

```{r}
productCat_list <- unique(productCat_basics$productName)
productCat_list <- gsub(" ", "_", productCat_list)
productCat_list <- gsub("\\(", "", productCat_list)
productCat_list <- gsub(")", "", productCat_list)
productCat_list <- gsub("\\|", "_", productCat_list)
productCat_list <- gsub("\\/", "_", productCat_list)

plot_product_metric_by_day <- function(metric) {
  ggplot(productCat_basics_spread, aes(x = date, color = "value")) +
    geom_line(aes_string(y = metric)) +
    theme_minimal() + xlab("") + ylab("value") +
    ggtitle(glue("28 day trend \n of {metric}")) +
    theme(plot.title = element_text(size = 10)) +
    scale_color_viridis(discrete = TRUE) + theme(legend.position = "none")
    
}

productplotlist <- list()
for (m in productCat_list) {
  productplotlist[[m]] <- plot_product_metric_by_day(m)
}

# arrange plots in a grid
do.call("grid.arrange", c(productplotlist, ncol=floor(sqrt(length(productplotlist)))))
```

Showing a summary and percentiles:

```{r}
skim_to_wide(productCat_basics_spread[2:ncol(productCat_basics_spread)]) %>%
  select(variable, mean, sd, p0, p50, p100, hist)
```


Checking E-commerce conversion metrics by channel Grouping to see how it looks.

```{r}
conversion_list <- unique(conversion_basics$channelGrouping)
conversion_list <- gsub(" ", "_", conversion_list)
conversion_list <- gsub("\\(", "", conversion_list)
conversion_list <- gsub(")", "", conversion_list)

plot_convmetric_by_day <- function(metric) {
  ggplot(conversion_basics_spread, aes(x = date, color = "value")) +
    geom_line(aes_string(y = metric)) +
    theme_minimal() + xlab("") + ylab("value") +
    ggtitle(glue("28 day trend of {metric}")) +
    theme(plot.title = element_text(size = 10)) +
    scale_color_viridis(discrete = TRUE) + theme(legend.position = "none")
    
}
conv_plotlist <- list()
for (m in conversion_list) {
  conv_plotlist[[m]] <- plot_convmetric_by_day(m)
}
# arrange plots in a grid
do.call("grid.arrange", c(conv_plotlist, ncol=floor(sqrt(length(conv_plotlist)))))
```


Showing a summary and percentiles:

```{r}
skim_to_wide(conversion_basics_spread[2:ncol(conversion_basics_spread)]) %>%
  select(variable, mean, sd, p0, p50, p100, hist)
```



Checking E-commerce conversion metrics by device category to see how it looks.

```{r}
conversiondevice_list <- unique(conversion_basics$deviceCategory)
conversiondevice_list <- gsub(" ", "_", conversiondevice_list)
conversiondevice_list <- gsub("\\(", "", conversiondevice_list)
conversiondevice_list <- gsub(")", "", conversiondevice_list)

plot_convdevicemetric_by_day <- function(metric) {
  ggplot(conversion_device_basics_spread, aes(x = date, color = "value")) +
    geom_line(aes_string(y = metric)) +
    theme_minimal() + xlab("") + ylab("value") +
    ggtitle(glue("28 day trend of {metric}")) +
    theme(plot.title = element_text(size = 10)) +
    scale_color_viridis(discrete = TRUE) + theme(legend.position = "none")
    
}

conv_device_plotlist <- list()
for (m in conversiondevice_list) {
  conv_device_plotlist[[m]] <- plot_convdevicemetric_by_day(m)
}

# arrange plots in a grid
do.call("grid.arrange", c(conv_device_plotlist, ncol=floor(sqrt(length(conv_device_plotlist)))))
```


Showing a summary and percentiles:

```{r}
skim_to_wide(conversion_device_basics_spread[2:ncol(conversion_device_basics_spread)]) %>%
  select(variable, mean, sd, p0, p50, p100, hist)
```

## Custom Dimensions and Metrics

`r customer` has configured `r ifelse(is.null(nrow(cmets$items)),0,nrow(cmets$items))` custom metrics and
`r ifelse(is.null(nrow(cdims$items)),0,nrow(cdims$items))` dimensions.

```{r} 
show_custom_vars <- function(x) {
  if (!is.null(nrow(x$items))) {
    type <- gsub('.*#custom', '', x$kind )
    df <- x$items[,c("index", "name", "scope", "active")] 
    
    # display table + add styling
    kable(df %>% 
            mutate(active = cell_spec(active, color = ifelse(!is.na(active), "green", "red"))),
          escape = FALSE,
          caption = glue("Custom {type} list for {customer}")) %>% 
      kable_styling(bootstrap_options = c("striped", "condensed", full_width=FALSE))
  }
}
show_custom_vars(cdims)
show_custom_vars(cmets)
```

### Custom Dim values

Here's a summary of the Custom Dimensions, and the number of values found:

```{r message=FALSE, warning = FALSE}
call_ga <- function(index = 1, n = 100, active = TRUE, name = "(no name)", id = viewid, dates = date_range) {
  
  # returns a list of 2:
  # summary = 1-row df with summary
  # results = df of results (top 10)
  
  ret <- list(
    summary = data.frame(
      index = index,
      name = name,
      active = active,
      rows_queried = 0,
      rows_found = 0
    ),
    results = data.frame()
  )
  
  if (!is.na(active) && active == TRUE) {
    df <- google_analytics(viewId = id, date_range = dates,
                           dimensions = c(glue("dimension{index}")),
                           metrics=c("users"),
                           order = order_type("users", sort_order = c("DESCENDING")),
                             max = n)
    ret <- list(
      summary = data.frame(
        index = index,
        name = name,
        active = active,
        rows_queried = n,
        rows_found = ifelse(is.null(nrow(df)),0,nrow(df))
      ),
      results = head(df,10)
    )
  }
  
  # return something
  ret
}
get_custom_vars <- function(x) {
  # input: cdims or cmets
  # output: list of 2 data frames:
  # 1: summary df
  # 2: value df of all dims/metrics
  
  dim_table <- data.frame()
  res_table <- data.frame()
      
  if (!is.null(nrow(x$items))) {
    type <- gsub('.*#custom', '', x$kind )
    df <- x$items[,c("index", "name", "scope", "active")] 
    
    if (type == "Dimensions") {
      # populate tables
      for (d in 1:nrow(df)) {
        c <- call_ga(index  = df[d,"index"],
                     active = df[d,"active"],
                     name   = df[d,"name"])
        dim_table <- rbind(dim_table, c$summary)
        
        if (length(c$results) > 0) {
          thistop10 <- c$results
          colnames(thistop10) <- c("dimension","users")
        } else {
          thistop10 <- data.frame(dimension = "(no result)", users = as.numeric(NA))
        }
        
        thistop10$name <- df[d,"name"]
        thistop10$index <- df[d,"index"]
        
        res_table <- rbind(res_table, thistop10)
      }
    } else {
      message("no custom metric support yet")
    }
  }
  
  list(summary_df = dim_table, content_df = res_table)
}
fetched_dims <- get_custom_vars(cdims)
# print table, add styling
kable(fetched_dims[["summary_df"]],
      escape = FALSE,
      caption = glue("Custom Dim summary for {customer}")) %>% 
  row_spec(which(fetched_dims[["summary_df"]]$rows_found == 0), color = "red") %>% 
  kable_styling(bootstrap_options = c("striped", "condensed"))
```

### Top 10 values per dimension

A long table, with the top 10 values of each dimension and the number of users per said value.

```{r}
df <- fetched_dims[["content_df"]] %>% select(index, name, dimension, users) %>% filter(!is.na(users))
kable(df,
      escape = FALSE,
      caption = glue("Custom Dim summary for {customer}")) %>% 
  row_spec(which(df$index %% 2 == 0), background = "#eeeeee") %>% 
  kable_styling(bootstrap_options = c("condensed"))
```